Concurrent Programming with Harmony

### 2020-07-28 book/paper.tex
Working on Chapter 10

What we want is a special kind of lock that allows either one writer
or one or more readers to be in the critical section.  This is called
a \emph{reader/writer lock}~\cite{CHP71}.
We will explore various ways of implementing reader/writer locks in
this chapter and future ones.

@@

While correct, it is not considered a good solution.
The solution is an example of what is called \emph{busy-waiting}
\index{busy waiting}%
(aka \emph{spin-waiting}):
\index{spin waiting}%
processes spin in a loop until some desirable application-level condition is met.
\autoref{fig:inc} also has an example of busy-waiting: the \texttt{main}
process waits for the other two processes to finish by checking their \texttt{done}
flags.

The astute reader might wonder if obtaining the mutex itself is an
example of busy-waiting.  After all, the Harmony \texttt{synch} implementation
of \texttt{lock()} spins in a loop until the mutex is available
(see \autoref{fig:spinlocks}).
However, there is a big difference.  As we pointed out, the processes that
are waiting for a spinlock are \emph{blocked}.
In most operating systems and programming language runtimes,
when a process acquires a mutex, the process is placed on a scheduling
queue and stops using CPU cycles until the mutex becomes available.
Harmony supports this with the \texttt{synchS} module.
Busy waiting disables all this: even when using the \texttt{synchS}
module, readers and writers only get suspended temporarily in case
there is contention for the mutex.
A process trying to obtain a read or write lock in
\autoref{fig:rwbusy} has to obtain the mutex repeatedly
until the read or write lock is available.

>>>>

What we want is a special kind of lock that allows either (i) one writer
or (ii) one or more readers to be in the critical section.  This is called
a \emph{reader/writer lock}~\cite{CHP71}.
We will explore various ways of implementing reader/writer locks in
this chapter and future ones.

@@

While correct, it is not considered a good solution.
The solution is an example of what is called \emph{busy-waiting}
\index{busy waiting}%
(aka \emph{spin-waiting}):
\index{spin waiting}%
processes spin in a loop until some desirable application-level condition is met,
wasting CPU cycles.
\autoref{fig:inc} also has an example of busy-waiting: the \texttt{main}
process waits for the other two processes to finish by checking their \texttt{done}
flags.

The astute reader might wonder if obtaining the mutex itself is an
example of busy-waiting.  After all, the Harmony \texttt{synch} implementation
of \texttt{lock()} spins in a loop until the mutex is available
(see \autoref{fig:spinlocks}).
However, there is a big difference.  As we pointed out, the processes that
are waiting for a spinlock are \emph{blocked}.
In most operating systems and programming language runtimes,
when a process acquires a mutex, the process is placed on a scheduling
queue and stops using CPU cycles until the mutex becomes available.
Harmony supports this with the \texttt{synchS} module.
Busy waiting disables all this: even when using the \texttt{synchS}
module, readers and writers only get suspended temporarily in case
there is contention for the mutex.
A process trying to obtain a read or write lock in
\autoref{fig:rwbusy} tries to obtain the mutex repeatedly.
